---
title: "Emperical anlaysis"
author: "Yoonsu Cho"
date: "23/04/2021"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

# Call the package
```{r set up}
library(MR.TRAM)
```

```r
y <- a
a$import(y)

a <- MultiAncestrySummarySet$new()
a$import(y)

```

# SBP-CHD
Exposures: SBP (EUR: systolic blood pressure; International Consortium of Blood Pressure, EAS: BBJ); 
Outcomes: CHD (EUR-EAS)
```{r}
a <- MultiAncestrySummarySet$new(
  exposure_ids=c("ieu-b-38", "bbj-a-52"), 
  outcome_ids=c("ieu-a-7", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
a$check_phenotypes(ids=a$exposure_ids)
a$check_phenotypes(ids=a$outcome_ids)

a$extract_instruments()
a$estimate_instrument_specificity(a$instrument_raw)

a$standardised_units(exp=a$instrument_raw)
a$estimate_instrument_specificity(a$standardised_exposure)

a$make_outcome_data(exp=a$instrument_raw, p_exp=1)
a$harmonised_dat(exp=a$instrument_raw, out=a$instrument_outcome)
a$perform_basic_sem()


a$make_outcome_data(exp=a$standardised_exposure, p_exp=1)
a$harmonised_dat(exp=a$standardised_exposure, out=a$instrument_outcome)
a$perform_basic_sem()


suppressMessages(a$extract_instrument_regions())
a$scan_regional_instruments()
a$plot_regional_instruments(instruments=a$instrument_maxz)

a$estimate_instrument_specificity(a$instrument_maxz)
a$standardised_units(exp=a$instrument_maxz)
a$estimate_instrument_specificity(a$standardised_exposure)

a$make_outcome_data(exp=a$standardised_exposure, p_exp=1)
a$harmonised_dat(exp=a$standardised_exposure, out=a$instrument_outcome)
a$perform_basic_sem()


```


Exposures: SBP (EUR: Systolic blood pressure, automated reading; MRC-IEU, EAS: BBJ); 
Outcomes: CHD (EUR-EAS)
```{r}
b <- MultiAncestrySummarySet$new(
  exposure_ids=c("ukb-b-19953", "bbj-a-52"), 
  outcome_ids=c("ieu-a-7", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
b$check_phenotypes(ids=b$exposure_ids)
b$check_phenotypes(ids=b$outcome_ids)

b$extract_instruments()
b$estimate_instrument_specificity(b$instrument_raw)

b$standardised_units(exp=b$instrument_raw)
b$estimate_instrument_specificity(b$standardised_exposure)

b$make_outcome_data(exp=b$instrument_raw, p_exp=1)
b$harmonised_dat(exp=b$instrument_raw, out=b$instrument_outcome)
b$perform_basic_sem()


b$make_outcome_data(exp=b$standardised_exposure, p_exp=1)
b$harmonised_dat(exp=b$standardised_exposure, out=b$instrument_outcome)
b$perform_basic_sem()


suppressMessages(b$extract_instrument_regions())
b$scan_regional_instruments()
b$plot_regional_instruments(instruments=b$instrument_maxz)

b$estimate_instrument_specificity(b$instrument_maxz)
b$standardised_units(exp=b$instrument_maxz)
b$estimate_instrument_specificity(b$standardised_exposure)

b$make_outcome_data(exp=b$standardised_exposure, p_exp=1)
b$harmonised_dat(exp=b$standardised_exposure, out=b$instrument_outcome)
b$perform_basic_sem()
```


Exposures: SBP (EAS: Systolic blood pressure, automated reading; Pan-UKB; EAS: BBJ)
Outcomes: CHD (EAS: Pan-UKB; EAS-BBJ)
```r
c <- MultiAncestrySummarySet$new(
  exposure_ids=c("ukb-e-4080_EAS", "bbj-a-52"), 
  outcome_ids=c("ukb-e-6150_p5_EAS", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
c$check_phenotypes(ids=c$exposure_ids)
c$check_phenotypes(ids=c$outcome_ids)
#Unit information is missing: Run x$standardise_units()
#Degree of agreement in instrument associations between NA and NA is NA
#Degree of agreement in instrument associations between  and  is 
c$extract_instruments()
c$check_instruments()
c$estimate_instrument_specificity(c$instrument_raw)
c$extract_instrument_regions()
c$scan_regional_instruments()
c$plot_regional_instruments(instruments=c$instrument_maxz)
c$estimate_instrument_specificity(c$instrument_maxz)
c$extract_outcome_data(exp=c$instrument_raw, standardise = T, p_exp=1)
c$perform_basic_sem()
c$extract_outcome_data(exp=c$instrument_maxz, standardise = T, p_exp=1)
c$perform_basic_sem()

```


Exposures: SBP (EUR: systolic blood pressure; International Consortium of Blood Pressure,EUR: EUR: Systolic blood pressure, automated reading); 
Outcomes: CHD (EUR-EAS)
```{r}
c <- MultiAncestrySummarySet$new(
  exposure_ids=c("ieu-b-38", "ukb-b-19953"), 
  outcome_ids=c("ieu-a-9", "ieu-a-7"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
c$extract_instruments()
c$estimate_instrument_specificity(c$instrument_raw)

c$standardised_units(exp=c$instrument_raw)
c$estimate_instrument_specificity(c$standardised_exposure)

c$make_outcome_data(exp=c$instrument_raw, p_exp=1)
c$harmonised_dat(exp=c$instrument_raw, out=c$instrument_outcome)
c$perform_basic_sem()

c$make_outcome_data(exp=c$standardised_exposure, p_exp=1)
c$harmonised_dat(exp=c$standardised_exposure, out=c$instrument_outcome)
c$perform_basic_sem()

suppressMessages(c$extract_instrument_regions())
c$scan_regional_instruments()
c$plot_regional_instruments(instruments=c$instrument_maxz)

c$estimate_instrument_specificity(c$instrument_maxz)
c$standardised_units(exp=c$instrument_maxz)
c$estimate_instrument_specificity(c$standardised_exposure)

c$make_outcome_data(exp=c$standardised_exposure, p_exp=1)
c$harmonised_dat(exp=c$standardised_exposure, out=c$instrument_outcome)
c$perform_basic_sem()
```
