---
title: "Emperical anlaysis"
author: "Yoonsu Cho"
date: "23/04/2021"
output: html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE
)
```

# Call the package
```{r set up}
library(MR.TRAM)
```

# SBP-CHD
Exposures: SBP (EUR: systolic blood pressure; International Consortium of Blood Pressure, EAS: BBJ); 
Outcomes: CHD (EUR-EAS)
```{r}
a <- MultiAncestrySummarySet$new(
  exposure_ids=c("ieu-b-38", "bbj-a-52"), 
  outcome_ids=c("ieu-a-7", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
a$extract_instruments()
a$check_instruments()
a$standardised_units()
a$estimate_instrument_specificity(a$instrument_raw)
suppressMessages(a$extract_instrument_regions())
a$scan_regional_instruments()
a$plot_regional_instruments(instruments=a$instrument_maxz)
a$estimate_instrument_specificity(a$instrument_maxz)
a$extract_outcome_data(exp=a$instrument_raw, standardise = F, p_exp=1)
a$perform_basic_sem()

a$extract_outcome_data(exp=a$instrument_maxz, standardise = F, p_exp=1)
a$perform_basic_sem()

```


Exposures: SBP (EUR: Systolic blood pressure, automated reading; MRC-IEU, EAS: BBJ); 
Outcomes: CHD (EUR-EAS)
```{r}
b <- MultiAncestrySummarySet$new(
  exposure_ids=c("ukb-b-19953", "bbj-a-52"), 
  outcome_ids=c("ieu-a-7", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
b$extract_instruments()
b$check_instruments()
b$estimate_instrument_specificity(b$instrument_raw)
suppressMessages(b$extract_instrument_regions())
b$scan_regional_instruments()
b$plot_regional_instruments(instruments=b$instrument_maxz)
b$estimate_instrument_specificity(b$instrument_maxz)
b$extract_outcome_data(exp=b$instrument_raw, p_exp=1)
b$perform_basic_sem()

b$extract_outcome_data(exp=b$instrument_maxz, p_exp=1)
b$perform_basic_sem()
```


Exposures: SBP (EAS: Systolic blood pressure, automated reading; Pan-UKB; EAS: BBJ)
Outcomes: CHD (EUR-EAS)
```r
c <- MultiAncestrySummarySet$new(
  exposure_ids=c("ukb-e-4080_EAS", "bbj-a-52"), 
  outcome_ids=c("ieu-a-7", "bbj-a-109"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
c$extract_instruments()
c$check_instruments()
c$estimate_instrument_specificity(c$instrument_raw)
c$extract_instrument_regions()
c$scan_regional_instruments()
c$plot_regional_instruments(instruments=c$instrument_maxz)
c$estimate_instrument_specificity(c$instrument_maxz)
c$extract_outcome_data(exp=c$instrument_raw, standardise = T, p_exp=1)
c$perform_basic_sem()
#lavaan WARNING:
#    the optimizer (NLMINB) claimed the model converged, but not all
#    elements of the gradient are (near) zero; the optimizer may not
#    have found a local solution use check.gradient = FALSE to skip
#    this check.lavaan WARNING: fit measures not available if model did not converge
c$extract_outcome_data(exp=c$instrument_maxz, standardise = T, p_exp=1)
c$perform_basic_sem()

```


Exposures: SBP (EUR: systolic blood pressure; International Consortium of Blood Pressure,EUR: EUR: Systolic blood pressure, automated reading); 
Outcomes: CHD (EUR-EAS)
```r
c <- MultiAncestrySummarySet$new(
  exposure_ids=c("ieu-b-38", "ukb-b-19953"), 
  outcome_ids=c("ieu-a-7", "ieu-a-7"), 
  bfiles=c("/work/yc16575/1kg_imp_gwasglue/EUR", "/work/yc16575/1kg_imp_gwasglue/EAS"), 
  plink = genetics.binaRies::get_plink_binary(), 
  radius=50000, 
  clump_pop="EAS"
)
c$extract_instruments()
c$check_instruments()
c$estimate_instrument_specificity(c$instrument_raw)
c$extract_instrument_regions()
c$scan_regional_instruments()
c$plot_regional_instruments(instruments=c$instrument_maxz)
c$estimate_instrument_specificity(c$instrument_maxz)
c$extract_outcome_data(exp=c$instrument_raw, p_exp=1)
c$perform_basic_sem()
#Error: attempt to use zero-length variable name

c$extract_outcome_data(exp=c$instrument_maxz, p_exp=1)
c$perform_basic_sem()
```
